Natural sources of vagueness and their implications

### 2017-05-16 book_chapter/paper.tex
Moved Section 2 back into the intro. Some other minor changes.

The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also need to be abandoned as a consequence.
One notion that seems to quickly be in peril is that of rationality.
In reference to philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
\begin{quote}
In the past, such philosophers have typically conjoined the claim that there is universal human agreement on the supreme desirability of truth with two further premises: that truth is correspondence to reality, and that reality has an intrinsic nature (that there is, in Nelson Goodman's terms, a Way the World Is).
[\ldots]
The rise of relatively democratic, relatively tolerant, societies in the last few hundred years is said to be due to the increased rationality of modern times, where `rationality' denotes the employment of an innate truth-oriented faculty.%
~\parencite*[1]{rorty_response_2000-1}
\end{quote}
By tying all of these notions tightly together, the fear could be that by dropping the picture of meaning as intimately tied to truth and logic we necessarily lose the ground on which rationality stands, and the whole edifice could collapse.
Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
Although humans are not necessarily consciously aware of the rules, they are considered to be nevertheless unconsciously following them when they act rationally.
This type of rationality is one that supposedly demarcates humans from other animals.

@@

We suggest that inspiration can be drawn from the later work of philosopher Ludwig Wittgenstein%
\footnote{To the extent that substantial views can be said to be defended by the author. Skipping over the debate (but see~\cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
His \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} adumbrate a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.

Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and reflect on their implications for how one thinks of rationality in the context of language use.
We start, in Section~\ref{sec:facets-rationality}, by clarifying some concepts that can help characterize different notions of rationality.
In Section~\ref{sec:signaling-games} we introduce the framework of signaling games and in Section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a special kind of signaling game, a so-called sim-max game.
Section~\ref{sec:finite-experience} reflects on the implications of these explanations for notions of rationality.
Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.

\section{Facets of rationality}
\label{sec:facets-rationality}

Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
Discussions around it, although touching on different aspects of the concept, often rely on a number of implicit assumptions.
We can start by considering what is deemed to be most important when characterizing the rationality of a given choice.
The aforementioned classical picture of language and meaning is focused on procedures.
Thus, a choice is deemed rational if it was arrived at by means of following proper rules of reasoning, like for example a logically valid deductive inference.
A \emph{procedural} account of rationality focuses on the means, rather than the ends.
One can also do the opposite and focus on the consequences of the choice instead.
\emph{Instrumental} rationality, as it is typically called\todo{references}, characterizes an agent's choice as rational if it maximizes the possibility of achieving a desired goal, regardless of the means.
The notion is linked to David Hume~\parencite*{hume_treatise_1738} and epitomized in his assertion that ``Reason is, and ought to only be the slave of the passions.''
This is close to a notion of rational choice that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility%
\footnote{\todo[inline]{discuss other options like minimax and such}}%
.
A more in-depth discussion of the opposition in the context of theoretical economics can be found, for example, in the work of Herbert A.~Simon~\parencite*[\emph{e.g.}][]{simon_rationality_1986}%
\footnote{Simon uses the term substantive instead of instrumental rationality, but the characterization is basically the same~\parencite[210-212]{simon_rationality_1986}.}%
.
%Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $EU(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.\todo{Brings in a particular model too soon}

When developing models where rationality is relevant, be it constructing a logical system or setting up a formal game, the assumed epistemic relation between whatever agents are involved and the elements of the model can come to bear on considerations of rationality.
Depending on how accurate and complete the agent's knowledge of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, etc, the verdict over the rationality of certain choices can vary.
An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
Models working within the classical picture typically do not make a distinction between modeler and agent, thus there is no room to express any epistemic gaps.
It follows that agents can only be assumed to be omniscient of the details involved.
By abstracting away from language users, these models cannot represent potential interactions between them, let alone repeated interaction and language change.
In other types of models, however, a further aspect of rational belief formation needs to be considered, namely the ability of agents to make accurate predictions about how other agents behave.
We can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate the actions and beliefs of other agents, and to predict medium/long term gains from repeated play.
%Strategic rationality is important for language use, because it is natural to assume that an agent's beliefs about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use said expression.\todo{too soon}

@@

%We draw attention to two additional aspects of rationality which will be helpful in the following.
Where the above definition of instrumental rationality is usually understood as requiring a single choice of act to maximize the expected utility in a single concrete decision situation, we might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
A choice mechanism is a general way of behaving for an agent involved in a variety of decision situations.
When considering only a single situation, rationality can only be \emph{local}.
If we take into account the possibility of the agent's choice or choice mechanism hinging on multiple situations, we can also talk about \emph{global} or \emph{ecological} rationality.
This is important because, hypothetically, there could be choice mechanisms that are sub-optimal at a local level (for each situation) but are actually perfectly rational at a global level.

@@

Note that we consider that, in theory, most combinations of these aspects are possible.
Although we have been pinning logic to the classical picture, this is only with the most traditional logical systems in mind.
Dynamic, epistemic, fuzzy, paraconsistent, and other types of logic\todo{cite Modal Logic for Open Minds?}, can potentially capture procedural rationality with limited non-strategic agents.
Game-theoretical models of language can, on the other hand, also assume local instrumental rationality with omnisciente highly strategic agents.
Our objective here is not to survey all the possibilities.
We want to focus on signaling games as a framework for the study of language use and meaning.
The vocabulary introduced in this section will, we hope, help inform the discussion that follows.

@@

The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like one would have in the classical picture.
However, we can already see how it attempts to address the study of meaning from a very different angle.
There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes \parencite[but see][]{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis}.
There is also no appeal to truth as a guide for our inquiry.
One advantage of such a paradigm shift is that we are no longer tied to the need of explaining how language hooks on to the world; this question is no longer of primary relevance.
We can merely focus on trying to see how agents can use signals to cope with the world and achieve their purposes.
On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
In the examples discussed so far, agents are assumed to possess and exercise rationality.
It is not, however, the procedural notion of rationality of the classical picture, but rather one that is instrumental and heavily strategic.

% Lewis signaling games can be seen as an alternative to the classical picture, even though the author himself might not have agreed with this.
% In order to get there, some aspects of the approach need to be refined.
% Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be a certain degree of required common knowledge between the players.
Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, strategic rationality, mutual ascription of strategic rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
Agents are thus envisioned as omniscient of the game and highly strategic.
These requirements can seem excessive, even more so if we consider how simple signaling systems are when compared to human languages.
The models were introduced in order to help explain how language could get off the ground as a conventional system without any sort of prior agreement.
However, if we consider the origins of language from a historical perspective, it seems implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages.
Furthermore, communication through simple message exchange is something that almost all animals do: monkeys use calls, birds use singing, bees use dances, ants use pheromone trails, and so on.
A plausible account of the origin of language should first explain how signaling systems like those could get started, without requiring high standards of rationality of the agents involved.

In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes to study signaling problems in evolutionary terms.
Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
Although the inspiration for this scenario is biological evolution, similar things could be said about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}, \emph{i.e.}~we can interpret these notions in terms of cultural evolution~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985}.
The principles can be captured in formal models that abstracts away from details of single interactions and individual agent's behavior, \emph{e.g.}~the replicator dynamics~\parencite{TaylorJonker1978:Evolutionary-St}.
The only thing relevant to the replicator dynamics are the relative proportions of strategies in a given population and the utility function.
Using it, one can compute which strategies evolve under which conditions.

Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalism to the descriptive side of the investigation.
Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
Focus is put on understanding how various ingredients of the model interact, and which results they produce, not on metaphysical concerns.
While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
This range of options goes hand in hand with a range of pictures of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of instrumental but limited and non-strategic rationality in the case of learning dynamics, to higher levels of rationality and even recursive strategic reasoning about the co-players' beliefs and choices.
Each of these can be utilized depending on the problem that one is interested in characterizing.
Thus, although Skyrms shows that rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where rationality can play a role.

The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under certain initial conditions and standard evolutionary pressures, and where it remains after small disturbances.
With these tools, not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which do not.
In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely. %; which one of the possible two will depend on their relative proportions in the original population.
More complex signaling problems may have different evolutionary outcomes, sometimes unexpected ones.
Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few.
More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \citet{FrankeWagner2014:Game-Theory-and}.

@@

% \subsection{Bounded rationality}
\textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can be boundedly rational, \emph{i.e.}~how vagueness could arise as a consequence of cost-saving limitations in the cognitive capacities of instrumentally rational agents.
The first proposal is called limited memory fictitious play (LMF) and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy.
Each agent plays an instrumentally rational best response (given the utility function) to their estimate of the other's strategy.
In order to observe the evolution of strategies in repeated interaction, the authors model several individual agents in actual play.
What the authors observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the characteristics of a vague language as characterized before.
However, each agent still commands a crisp language, which is inadequate if the intention is to capture how vagueness presents itself in human languages.

@@

But why should agents evolve to generalize?
%First, it is not argued that vagueness in itself is beneficial, but that it is a by-product of a learning mechanism.
O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages, though less obvious ones (see \textcite{oconnor_evolving_2015} for the detailed argument).
Considerations of optimality of strategies, for example by considering the evolutionary stable states, are typically made in terms of hypothetical atemporal comparisons of expected utility.
However, GRL has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a smaller period of time.
In a naturalized evolutionary setting, where we care about bounded rationality in a multi-context environment, learning speed is an advantage which should also be taken into account.
Imagine an initial population of agents with random strategies, some using GRL and others using classical HRL to adapt to each other.
Although the latter type of agent can hypothetically develop a precise and more efficient signaling system, agents using GRL would coordinate on vague signaling strategies with high (though not optimal) expected utility sooner than agents using HRL.
In such a scenario, they could drive the other agents to extinction before the latter had time to achieve coordination and reap the benefits of a more precise signaling system.

@@

All of these observations support the weakening of our modeling assumptions.

>>>>

The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also need to be abandoned as a consequence.
One notion that seems to quickly be in peril is that of rationality.
In reference to philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
\begin{quote}
In the past, such philosophers have typically conjoined the claim that there is universal human agreement on the supreme desirability of truth with two further premises: that truth is correspondence to reality, and that reality has an intrinsic nature (that there is, in Nelson Goodman's terms, a Way the World Is).
[\ldots]
The rise of relatively democratic, relatively tolerant, societies in the last few hundred years is said to be due to the increased rationality of modern times, where `rationality' denotes the employment of an innate truth-oriented faculty.%
~\parencite*[1]{rorty_response_2000-1}
\end{quote}
By tying all of these notions tightly together, the fear could be that by dropping the picture of meaning as intimately tied to truth and logic we necessarily lose the ground on which rationality stands, and the whole edifice could collapse.
Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions. % (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
Although humans are not necessarily consciously aware of the rules, they are considered to be nevertheless unconsciously following them when they act rationally.
This type of rationality is one that supposedly demarcates humans from other animals.

@@

We suggest that inspiration can be drawn from the later work of philosopher Ludwig Wittgenstein%
\footnote{To the extent that substantial views can be said to be defended by the author. Skipping over the debate (but see~\cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
His \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} adumbrate a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.

Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and reflect on their implications for how one thinks of rationality in the context of language use.
Before we start, we want establish some vocabulary that will be useful later on.
%
%
%\section{Facets of rationality}
%\label{sec:facets-rationality}
%
Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
Discussions around it, although touching on different aspects of the concept, often fail to clearly demarcate them.
We can start by considering what is deemed to be most important when characterizing the rationality of a given choice.
The aforementioned classical picture of language and meaning is focused on dichotomies like true and false, meaningful and meaningless, correct and incorrect.
A sentence can be true or false only if it is meaningful, and it is meaningful if it is constructed according to correct rules.
Rationality is intimately connected with the ability to follow good procedures, not only of sentence production, but ultimately of sentence combination and reasoning (think of what is required for making a logically valid deductive inference).
A \emph{procedural} account of rationality thus focuses on the means, rather than the ends.
One can also do the opposite and focus on the consequences instead.
\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if it maximizes the possibility of achieving a desired goal, regardless of the means.
The notion is linked to David Hume~\parencite*{hume_treatise_1738}, epitomized in his assertion that ``Reason is, and ought to only be the slave of the passions.''
Instrumental rationality is close to a notion of rational choice that is used in economics and game theory%
\footnote{In fact, it has been argued~\parencite{vanderschraaf_informal_1998} that Hume's whole account of convention is very closely in line with modern game theory.}%
: agents are rational if and only if they make decisions that maximize their expected utility.
A more in-depth discussion of the opposition in the context of theoretical economics can be found, for example, in the work of Herbert A.~Simon~\parencite*[\emph{e.g.}][]{simon_rationality_1986}%
\footnote{Simon uses the term substantive instead of instrumental rationality, but the characterization is basically the same~\parencite[210-212]{simon_rationality_1986}.}%
.
%Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $EU(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.\todo{Brings in a particular model too soon}
% The dichotomy between procedural and instrumental views of rationality is not an exclusive one, since usually both means and ends are taking into account, but it helps us describe where the emphasis is placed on.

When developing models where rationality is relevant, be it constructing a logical system or setting up a formal game, the assumed epistemic relation between whatever agents are involved and the elements of the model can come to bear on considerations of rationality.
Depending on how accurate and complete the agent's knowledge of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, etc, the verdict over the rationality of certain choices can vary.
An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
Models working within the classical picture typically do not make a distinction between modeler and agent, thus there is no room to express any epistemic gaps.
It follows that agents can only be assumed to be omniscient of the details involved.
By abstracting away from language users, these models cannot represent potential interactions between them, let alone repeated interaction and language change.
In other types of models, however, a further aspect of rational belief formation needs to be considered, namely the ability of agents to make accurate predictions about how other agents behave.
We can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate the actions and beliefs of other agents, and to predict medium/long term gains from repeated play.
%Strategic rationality is important for language use, because it is natural to assume that an agent's beliefs about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use said expression.\todo{too soon}
Lack of perfect knowledge of the situation or lack of strategicness can be caused by many possible factors.
These include, but are not limited to, limitations in handling information (receiving, storing, retrieving, transmitting) and limited computational resources to solve complex problems.
We can talk about \emph{bounded} rationality to characterize the choices of such agents.

@@

%We draw attention to two additional aspects of rationality which will be helpful in the following.
Where the above definition of instrumental rationality is usually understood as requiring a single choice of act to maximize the expected utility in a single concrete decision situation, we might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
A choice mechanism is a general way of behaving for an agent involved in a variety of decision situations.
When considering only a single situation, rationality can only be \emph{local}.
If we take into account the possibility of the agent's choice or choice mechanism hinging on multiple situations, we can also talk about \emph{global} or \emph{ecological} rationality.
This can be important because, hypothetically, there could be choice mechanisms that are sub-optimal at a local level (for each situation) but are actually perfectly rational at a global level.

@@

Note that we consider that, in theory, most combinations of these aspects are possible.
Although we have been pinning logic to the classical picture, this is only with the most traditional logical systems in mind.
We are not denying that advances in dynamic, epistemic, fuzzy, paraconsistent, and other types of logic could potentially enable one to capture procedural rationality with different characteristics.
Game-theoretical models of language can, on the other hand, also assume local instrumental rationality with omniscient highly strategic agents.
Our objective here is not to survey all the possibilities.
We want to focus on signaling games as a framework for the study of language use and meaning.
The vocabulary just introduced will, we hope, help inform the discussion that follows.
We proceed by introducing the framework of signaling games in Section~\ref{sec:signaling-games}.
In Section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a particular kind of signaling game.
Section~\ref{sec:finite-experience} reflects on the implications of these explanations for considerations of rationality.
Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.

@@

The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like one would have in the classical picture.
However, we can already see how it attempts to address the study of meaning from a very different angle.
There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific goals~\parencite[but see][]{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis}.
There is also no appeal to truth as a guide for our inquiry.
One advantage of such a paradigm shift is that we are no longer tied to the need of explaining how language hooks on to the world; this question is no longer of primary relevance.
We can merely focus on trying to see how agents can use signals to cope with the world and achieve their goals.
On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
In the examples discussed so far, agents are assumed to possess and exercise rationality.
Unlike in the classical picture, it is an instrumental, rather than procedural, rationality, but one where agents are still omniscient and highly strategic.

% Lewis signaling games can be seen as an alternative to the classical picture, even though the author himself might not have agreed with this.
% In order to get there, some aspects of the approach need to be refined.
% Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be a certain degree of required common knowledge between the players.
Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, strategic rationality, mutual ascription of strategic rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
Agents are thus envisioned as omniscient of the game and highly strategic.
These requirements can seem excessive, even more so if we consider how simple signaling systems are when compared to human languages.
The models were introduced in order to help explain how language could get off the ground as a conventional system without any sort of prior agreement.
However, if we consider the origins of language from a historical perspective, it seems implausible that the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages possessed such advanced rationality.
Furthermore, communication through simple message exchange is something that almost all animals do: monkeys use calls, birds use singing, bees use dances, ants use pheromone trails, and so on.
A plausible account of the origin of language should first explain how signaling systems like those could get started, without requiring high standards of rationality of the agents involved.

In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes to study signaling problems in evolutionary terms.
Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
Although the inspiration for this scenario is biological evolution, similar things could be said about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}, \emph{i.e.}~we can interpret these notions in terms of cultural evolution~\parencite{dawkins_selfish_1978,boyd_culture_1985}.
The principles can be captured in formal models that abstract away from details of single interactions and behavior of individual agents, for example in the replicator dynamics~\parencite{TaylorJonker1978:Evolutionary-St}.
The only thing relevant to the replicator dynamics are the relative proportions of strategies in a given population and the utility function.
Using it, one can compute which strategies evolve under which conditions.

Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalism to the descriptive side of the investigation.
Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
The shift in perspective allows interpretations that accommodate for limited non-strategic agents.
% Focus is put on understanding how various ingredients of the model interact, and which results they produce, not on metaphysical concerns.
While the general framework manages to abstract quite some details away from the formalization, it nevertheless leaves room for them, especially when it comes to the dynamics.
We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
This range of options goes hand in hand with a range of pictures of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of instrumental but limited and non-strategic rationality in the case of learning dynamics, to higher levels of rationality and even recursive strategic reasoning about the co-players' beliefs and choices.
Each of these can be utilized depending on the problem that one is interested in characterizing.
Thus, although Skyrms shows that high requirements of rationality are not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions between highly strategic agents.

The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under certain initial conditions and standard evolutionary pressures, and where it remains after small disturbances.
With these tools, we can better understand why signaling systems are stable even without any strong assumptions of rationality.
We can also map out which initial conditions drive the system towards which equilibria and which do not.
In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely. %; which one of the possible two will depend on their relative proportions in the original population.
More complex signaling problems may have different evolutionary outcomes, sometimes unexpected ones.
Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few.
More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \citet{FrankeWagner2014:Game-Theory-and}.

@@

% \subsection{Bounded rationality}
\textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can be boundedly rational, \emph{i.e.}~how vagueness could arise as a consequence of cost-saving limitations in the cognitive capacities of instrumentally rational agents.
The first proposal is called limited memory fictitious play (LMF) and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy.
Each agent plays an instrumentally rational best response (given the utility function) to their estimate of the other's strategy.
In order to study the evolution of strategies in repeated interaction, the authors model several individual agents in actual play.
What they observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the characteristics of a vague language as characterized before.
However, each agent still commands a crisp language, which is inadequate if the intention is to capture how vagueness presents itself in human languages.

@@

But why should agents evolve to generalize?
%First, it is not argued that vagueness in itself is beneficial, but that it is a by-product of a learning mechanism.
O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages, though less obvious ones (see \textcite{oconnor_evolving_2015} for the detailed argument).
Considerations of optimality of strategies, for example by considering the evolutionary stable states, are typically made in terms of hypothetical atemporal comparisons of expected utility.
However, GRL has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a shorter period of time.
In a naturalized evolutionary setting, where we care about bounded rationality in a multi-context environment, learning speed is an advantage which should also be taken into account.
Imagine an initial population of agents with random strategies, some using GRL and others using classical HRL to adapt to each other.
Although the latter type of agent can hypothetically develop a precise and more efficient signaling system, agents using GRL would coordinate on vague signaling strategies with high (though not optimal) expected utility sooner than agents using HRL.
In such a scenario, they could drive the other agents to extinction before the latter had time to achieve coordination and reap the benefits of a more precise signaling system.

@@

All of these observations support the weakening of the modeling assumptions.
